---
title: "x3p basics"
author: "Heike Hofmann"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## The file format 

The x3p file format is an xml based file format created to describe digital surface measurements.  x3p has been developed by OpenFMC (Open Forensic Metrology Consortium, see http://www.openfmc.org/) and has been adopted as ISO ISO5436 – 2000.
x3p files are a zip archive of a directory consisting of an xml file of meta information and a matrix of numeric surface measurements. 

## Functionality

`x3ptools` implements basic functionality for `x3p` files: 

- reading with `read_x3p`
- writing files with `write_x3p`
- displaying files with `image_x3p`
- rotating and transposing images: `rotate_x3p` and `transpose_x3p`
- changing meta information using templates with `addtemplate_x3p`
- casting between an x3p format and a data frame: `x3p_tp_df` and `df_to_x3p`
- sample from an x3p file with `sample_x3p`


## x3p objects

Internally, x3p objects are stored as a list consisting of the surface matrix (the measurements) and meta information in four records: header info, feature info, general info, and matrix info:

```{r}
library(x3ptools)
logo <- read_x3p(system.file("csafe-logo.x3p", package="x3ptools"))
names(logo)
```

The four info objects specify the information for Record1 through Record4 in the xml file. An example for an xml file is provided with the package and can be accessed as 
`system.file("templateXML.xml", package="x3ptools")`. 

`header.info`  contains the information relevant to interpret locations for the surface matrix:

```{r}
logo$header.info
```
`matrix.info` expands on `header.info` and provides the link to the surface measurements in binary format.

`general.info` consists of information on how the data was captured, i.e. both author and capturing device are specified here.

`feature.info` is informed by the header info and provides the structure for storing the information.

While these pieces can be changed and adapted manually, it is more convenient to save information on the capturing device and the creator in a separate template and bind measurements and meta information together with the command `addtemplate_x3p`.



## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
